---
title: "Introduction to Water Detection"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to Water Detection}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

**This Vignette was produced using RMarkdown documentation, you can find most of the syntax here: https://www.markdownguide.org/cheat-sheet/** <br />
<br />
**This package was built using library(devtools) for package building, library(roxygen2) for documentation, and library(pkgdown) for the generation of docs folder**
<br />
<br />
Welcome to the Introduction of the "WaterDetection" package. In this Vignette I will run you through the necessary steps to use this package and produce time-series graphs that represent the amount of water in a group of Zones of Interest (ZOI).
<br />


## Main Setup and Data
### **1** - Install and load package
The first step would obviously be to install and load our package, "WaterDetection". Use Packages > Install from file and use the file (.tar.gz probably) provided. Then load the package:

```{r setup}
library(WaterDetection)
```

or use this line, that will work as soon as our package is available on RCAN:

```{r, eval=FALSE}
if (!require('WaterDetection')) install.packages('WaterDetection'); library ('WaterDetection')
```
### **2** - Read Composite Input Data
We will need data, and as such you should define variables (lists) with your satellite imagery data. In this package, we'd like to have a list of file paths. The path used should point to a folder where you can find **pre-processed monthly composites** that correspond to a single year. This should be done **once for each year of data** (you should organize data 1 folder = 12 months of data) you'd like to use. Make sure to include **full.names = TRUE**. The usage of .VRT has not been tested yet, and as such tile-by-tile processing is recommended. For one year, Example:

```{r, eval=FALSE}
CompositeListYearX = list.files(path = "Whatever_File_Path", pattern = "Whatever_Identifying_Pattern", full.names = TRUE)
```

Or, for a set of 6 years (2017, 2018, 2019, 2020, 2021, 2022):

```{r, eval=FALSE}
CompositeListYear2017 = list.files(path = "Whatever_File_Path", pattern = "Whatever_Identifying_Pattern", full.names = TRUE)
CompositeListYear2018 = list.files(path = "Whatever_File_Path", pattern = "Whatever_Identifying_Pattern", full.names = TRUE)
CompositeListYear2019 = list.files(path = "Whatever_File_Path", pattern = "Whatever_Identifying_Pattern", full.names = TRUE)
CompositeListYear2020 = list.files(path = "Whatever_File_Path", pattern = "Whatever_Identifying_Pattern", full.names = TRUE)
CompositeListYear2021 = list.files(path = "Whatever_File_Path", pattern = "Whatever_Identifying_Pattern", full.names = TRUE)
CompositeListYear2022 = list.files(path = "Whatever_File_Path", pattern = "Whatever_Identifying_Pattern", full.names = TRUE)
```
### **3** - Read NDVI Input Data
This algorithm also uses the NDVI. If you already have NDVI calculated for each month of your data, you should now define lists that contain the file paths for NDVI. Otherwise you need to calculate it. This should be done **once for each year of data** (you should organize data 1 folder = 12 months of data)  you'd like to use. Make sure to include **full.names = TRUE**. For one year, again:

```{r, eval=FALSE}
NDVIListYearX = list.files(path = "Whatever_File_Path", pattern = "Whatever_Identifying_Pattern", full.names = TRUE)
```

Or, for a set of 6 years (2017, 2018, 2019, 2020, 2021, 2022):

```{r, eval=FALSE}
NDVIListYear2017 = list.files(path = "Whatever_File_Path", pattern = "Whatever_Identifying_Pattern", full.names = TRUE)
NDVIListYear2018 = list.files(path = "Whatever_File_Path", pattern = "Whatever_Identifying_Pattern", full.names = TRUE)
NDVIListYear2019 = list.files(path = "Whatever_File_Path", pattern = "Whatever_Identifying_Pattern", full.names = TRUE)
NDVIListYear2020 = list.files(path = "Whatever_File_Path", pattern = "Whatever_Identifying_Pattern", full.names = TRUE)
NDVIListYear2021 = list.files(path = "Whatever_File_Path", pattern = "Whatever_Identifying_Pattern", full.names = TRUE)
NDVIListYear2022 = list.files(path = "Whatever_File_Path", pattern = "Whatever_Identifying_Pattern", full.names = TRUE)
```
### **4** and **5** - Define a Shapefile (Path and Name) for our ZOI
Now it would take ages and it would be counter-productive (and also not very accurate) to apply this algorithm to the entirety of our data. As such, it is recommended that you produce a **shapefile (.shp)** with **ZOI (Zones of Interest)**. These zones should be zones where water is expected to be (say zones defining damns, lakes). These zones can (and probably should) come from the most recent mapping you have (example: COS2018, class 9.1.2.X). Then, define a variable with a path to that shapefile, and another variable with the shapefile name. **These variables should be global, if possible (using <<- instead of <-). However, caution using global variables regarding your overall work.** Example:

```{r, eval=FALSE}
ZOI_Shapefile_Path <<- "Whatever_Shapefile_Path"
Shapefile_Name <<- "Whatever_Shapefile_Name" # Example: COS2018_912X_Planos_de_agua_UTM.shp
```
### **6** - Define a list of Years to calculate
Define a list of Years with the same years you want to calculate;
Now we should define a variable with a list of years that we will calculate. This will help other functions work properly. Other solutions might automate this but it will severely complicate the code without any real need. Let's say you're using 6 years of data as per previous example. **This variable should be global, if possible (using <<- instead of <-). However, caution using global variables regarding your overall work.** Then:

```{r, eval=FALSE}
Year_List <<- list(2017,2018,2019,2020,2021,2022)
```
### **7** - Define a general Output path
Now let's define a path for the general output of our Algorithm. This path will eventually contain all the images for all ZOIs from every step of the process. Let's say:

```{r, eval=FALSE}
Output_Path <<- "Whatever_Path" #MAXIMUM CHARACTER SIZE 118
```

### **8** and **9** - Define thresholds (NDVI and RGBSum)
This algorithm uses both NDVI and a sum of RGB bands to detect water. We should now define what thresholds it should use to differentiate water from non-water. Define two variables, one for **NDVI_Threshold** and one for **RGBSum_Threshold**. I recommend 0 and 800, respectively. These variables should be a list with just one element. This was a preparation for further automation that wasn't completed (multiple thresholds per run). As such:

```{r, eval=FALSE}
NDVI_Threshold <<- list(0)
RGBSum_Threshold <<- list(800)
```

### **10** - Create Basic Output Folders
Although not perfectly automated, there is still a severe amount of automation in this package. We can now automate the creation of output folders in the output path we created in step 7. Using the function **CreateOutputFolders()** with the arguments **Year_List** from step 6 and **OutputPath** from step 7. Keep in mind this function has no return.

```{r, eval=FALSE}
CreateOutputFoldersALL(Year_List, Output_Path)
```

### **11** - Define variables in memory containing export paths
This package uses a lot of different export paths, and thus they need to be defined. This process is automated with the use of **GenerateExportPaths()**. You should have an export path for all these 13 different things **(a set for each year, XXXX = Year)**:
<br />
<ol>
1. ExportPath_NDVI_All_XXXX
2. ExportPath_Composite_All_XXXX
3. ExportPath_RGBSum_All_XXXX
4. ExportPath_NDVI_Binary_All_XXXX
5. ExportPath_RGBSum_Binary_All_XXXX
6. ExportPath_Algorithm_All_XXXX
7. ExportPath_NDVI_Individual_XXXX
8. ExportPath_Composite_Individual_XXXX
9. ExportPath_RGBSum_Individual_XXXX
10. ExportPath_NDVI_Binary_Individual_XXXX
11. ExportPath_RGBSum_Binary_Individual_XXXX
12. ExportPath_Algorithm_Individual_XXXX
13. ExportPath_CSV_XXXX
<br />
<br />
**And three more paths (not repeated by year):**
<br />
<br />
14. ExportPath_Graphs_Individual
15. ExportPath_Graphs_Overlap_Individual
16. ExportPath_Shapefiles_Individual
</ol>
<br />
Simply use as argument the year list from step 6 and the output path from step 7:
```{r, eval=FALSE}
GenerateExportPaths(Year_List, Output_Path)
```
### **12** - Cut our ZOI shapefile into Individual Shapefiles and Export Them
Now let's use **CropShapefile_ExportToPath()** to cut the Shapefile that contains our ZOI into diferent shapefiles, one for each ZOI. This is useful for later processing and also to build maps in GIS software (since we now have the boundaries of each ZOI). Note that I used **ZOI_Shapefile_Path**, **Shapefile_Name** and **ExportPath_Shapefiles_Individual** from steps 4, 5 and 11. This function will return a list of individual shapefile objects (ordered by OBJECT_ID from the atribute table), which we will store in a variable **ZOI_Shapefile_List**.
```{r, eval=FALSE}
ZOI_Shapefile_List = CropShapefile_ExportToPath(ZOI_Shapefile_Path, Shapefile_Name, ExportPath_Shapefiles_Individual)
```
### **13** - Build Vectors using defined thresholds (to reclassify rasters later)
Then you should build some vectors that will be used to reclassify rasters later. Use **VectorBuilder()** with the values defined in step 8 and 9. Remenber to define lists in steps 8 and 9 and then call the first element of each list using [[1]]:
```{r, eval=FALSE}
VectorBuilder(NDVI_Threshold[[1]], RGBSum_Threshold[[1]])
```
# Processing Steps by Year

I will now give you a run down of steps to take for each year of data you use. Later, all of the results of these processes will be put together to help us build time-series graphs for each ZOI.

## Processing NDVI

### **14** - Build a list of NDVI Bricks 
R works best when you turn your raster data into workable "stacks" or "bricks". In this case, we will use **BuildBrickList()** to transform our **NDVIListYearX** into workable "bricks". Simply:
```{r, eval=FALSE}
NDVI_YearX_BL = BuildBrickList(NDVIListYearX)
```
### **15** - Crop And Mask our NDVI imagery to the ZOI in our Shapefile
Now we can the function CropAndMask() to, well, Crop and Mask our data to the areas defined by our Shapefile. This will severely reduce processing time, as areas where Water is expected are a relatively low percentage of the total area. Note that I used **NDVI_YearX_BL** from step 14. Example:
```{r, eval=FALSE}
NDVI_YearX_ZOI = CropAndMask(NDVI_YearX_BL, Shapefile_Path, Shapefile_Name)
```
### **16** - Export the Cropped and Masked NDVI
We now have NDVI, relating to Year X, exclusively inside the ZOIs. We can use **ExportRasters()** to save those rasters, just in case we need them for some sort of analysis. Please note I used as arguments the NDVI to be exported, the ExportPath for NDVI_All (as produced in step 11), a string for a general file name (month name will be appended in the end so please end the name with _), a string to print as each NDVI is exported and a string to print in the end. These strings are useful if you ever need to debug something. As such:

```{r, eval=FALSE}
ExportRasterList(NDVI_YearX_ZOI, ExportPath_NDVI_All_YearX, "NDVI_All_YearX_", "Finished Exporting NDVI ", "Finished Exporting all NDVI")
```

### **17** - Create folders for the NDVI of each Individual ZOI
We now need to create individual folders for each ZOI. These folders will store the Individual NDVI, organized by ZOI. We can use **CreateIndividualFolders()**, with
**ExportPath_NDVI_Individual_YearX** (as per step 11) and **ID_list** (global variable produced with step 12 that represents a list of OBJECT_ID, one for each ZOI). That being said:
```{r, eval=FALSE}
CreateIndividualFolders(ExportPath_NDVI_Individual_YearX, ID_list)
```
### **18** - Cut and Export Individual ZOI NDVI and store those in a matrix in memory
We can now cut and export the monthly NDVI for each ZOI and store them in a matrix. Matrix[[1]] is a list of monthly NDVI for ZOI 1 and Matrix[[1]][[1]]
  is the the first month of the agricultural Year X for ZOI 1 (assuming entry data is 12 months and sorted by Agricultural Year). Use **Individual_Product()** with arguments: **NDVI_YearX_ZOI** from step 18, **ZOI_Shapefile_List** from step 12, **ExportPath_NDVI_Individual_YearX** from step 11, ID_list (global variable) from step 12, a name for each file (string "XXX_", as the function will attach the ZOI ID in the end), a string to print as each NDVI is exported, a string to print as each ZOI is completed and a string to print in the end;
```{r, eval=FALSE}
NDVI_YearX_IndividualZOI_ListOfLists = IndividualZOIProduct(NDVI_YearX_ZOI, ZOI_Shapefile_List, ExportPath_NDVI_Individual_YearX, ID_list, "NDVI_YearX_ZOI_", "Singular NDVI Exported ", "Completed ZOI ", "Completed All ZOI")
``` 
### **19** - Reclassify the NDVI with NDVI_Threshold (not the individual ZOI matrix)
Although we have the normal NDVI, we will need to reclassify it using our **NDVI_Threshold** from step 8. This, in conjunction with a reclassified RGBSum (to be done later) will produce our Water Detection algorithm. For this matter, you can use **ReclassRasterList()**:
```{r, eval=FALSE}
NDVI_YearX_ZOI_Binary = ReclassRasterList(NDVI_YearX_ZOI, VectorNDVI, "Reclassfied NDVI ", "Reclassified all NDVI")
```
### **20** - Export the recently reclassified NDVI
Better to export it just in case we need it later for some other purpose.
```{r, eval=FALSE}
ExportRasterList(NDVI_YearX_ZOI_Binary, ExportPath_NDVI_Binary_All_YearX, "Binary_NDVI_All_YearX_", "Finished Exporting Binary NDVI ", "Finished Exporting all Binary NDVI")
```
### **21** - Create folders for the Binary NDVI of each Individual ZOI
Let's create some more folders for each ZOI, named after the threshold selected in step 8, using **CreateIndividualFolders2()** this time, since it takes as input an additional argument:
```{r, eval=FALSE}
CreateIndividualFolders2(ExportPath_NDVI_Binary_Individual_YearX, ID_list, NDVI_Threshold)
```
### **22** - Cut and Export Individual ZOI Binary NDVI and store those in a matrix in memory
We can now cut and export the monthly Binary_NDVI for each ZOI and store them in a matrix. Matrix[[1]] is a list of monthly Binary NDVI for ZOI 1 and Matrix[[1]][[1]] is the the first month of the agricultural Year X for ZOI 1 (assuming entry data is 12 months and sorted by Agricultural Year). Use **Individual_Product2()** with arguments: **NDVI_YearX_ZOI_Binary** from step 18, **ZOI_Shapefile_List** from step 12, **ExportPath_NDVI_Binary_Individual_YearX** from step 11, ID_list (global variable) from step 12, a name for each file (string "XXX_", as the function will attach the ZOI ID in the end), a string to print as each Bianry NDVI is exported, a string to print as each ZOI is completed and a string to print in the end;
```{r, eval=FALSE}
NDVI_Binary_YearX_IndividualZOI_ListofLists = IndividualZOIProduct2(NDVI_YearX_ZOI_Binary, ZOI_Shapefile_List, ExportPath_NDVI_Binary_Individual_YearX, ID_list, NDVI_Threshold[[1]], "Binary_NDVI_YearX_ZOI_", "Singular Binary NDVI Exported ", "Completed ZOI ", "Completed All ZOI"))
``` 
## Processing RGBSum 
### **23** - Build Composite Bricks
This is a similar process to NDVI. First let's start by building Bricks from our data using **BuildBrickList()** and **CompositeListYearX** from step 2:
```{r, eval=FALSE}
Composite_YearX_BL = BuildBrickList(CompositeListYearX)
```
### **24** - Crop And Mask our Composite imagery to the ZOI in our Shapefile
Let's now Crop and Mask our data to our original Shapefile using **CropAndMask()**, **ZOI_Shapefile_Path** and **Shapefile_Name** from step 2:
```{r, eval=FALSE}
Composite_YearX_ZOI = CropAndMask(Composite_YearX_BL, Shapefile_Path, Shapefile_Name)
```
### **25** - Export the Cropped and Masked Composites
Export the result of CropAndMask, using **ExportPath()** and **ExportPath_Composite_All_YearX**, just in case we need later. Remeber to add a name for the file (XXX_, as the function will attach month name to the end, assuming data is 12 month and sorted by agricultural year), a string to print after each month is exported and a string to print in the end:
```{r, eval=FALSE}
ExportRasterList(Composite_YearX_ZOI, ExportPath_Composite_All_YearX, "Composite_All_YearX_", "Finished Exporting Composite ", "Finished Exporting all Composites")
```
### **26** - Create folders for the Composite of each Individual ZOI
Let's make sure we have individual folders for the Individual Composites by running **CreateIndividualFolders()** using **ExportPath_Composite_Individual_YearX** and **ID_list**: 
```{r, eval=FALSE}
CreateIndividualFolders(ExportPath_Composite_Individual_YearX, ID_list)
```
### **27** - Cut and Export Individual ZOI Composites and store those in a matrix in memory
Now let's again cut and export the monthly Composites for each ZOI and store them in a matrix where Matrix[[1]] is a list of monthly Composite for ZOI 1 and Matrix[[1]][[1]] is the the first month of the agricultural Year X for ZOI 1 (assuming entry data is 12 months and sorted by Agricultural Year) using as arguments: **Composite_YearX_ZOI** from step 24, **ZOI_Shapefile_List** from step 15, **ExportPath_Composite_Individual_YearX** from step 11, **ID_list** (global variable that should be created from running step 12), a string name for the files (XXX_, as the function will attach ZOI ID and month to the end), a string to print as each composite is exported, a string to print as each ZOI is finished and a string to print in the end;
```{r, eval=FALSE}
Composite_YearX_IndividualZOI_ListOfLists = IndividualZOIProduct(Composite_YearX_ZOI, ZOI_Shapefile_List, ExportPath_Composite_Individual_YearX, ID_list, "Composite_YearX_ZOI_", "Singular Composite Exported ", "Completed ZOI ", "Completed All ZOI")
```
### **28** - Calculate a sum of the RGB bands on our Composites (not the individual matrix)
But composites aren't useful without any processing, so we need to calculate a sum of the RGB bands. Just use **CalculateRGBSum()** with our **Composite_YearX_ZOI**:
```{r, eval=FALSE}
RGBSum_YearX_ZOI = CalculateRGBSum(Composite_YearX_ZOI)
```
### **29** - Export the recently calculated sum of RGB bands
Export the result, just in case we need it later, using **ExportRasterList()** with **ExportPath_RGBSum_All_YearX**:
```{r, eval=FALSE}
ExportRasterList(RGBSum_YearX_ZOI, ExportPath_RGBSum_All_YearX, "RGBSum_All_YearX_", "Finished Exporting RGBSum ", "Finished Exporting all RGBSum")
```
### **30** - Create folders for the RGBSum of each Individual ZOI
Let's make sure we have individual folders for the Individual RGBSum by running **CreateIndividualFolders()** using **ExportPath_RGBSum_Individual_YearX** and **ID_list**: 
```{r, eval=FALSE}
CreateIndividualFolders(ExportPath_RGBSum_Individual_YearX, ID_list)
```
### **31** - Cut and Export Individual ZOI RGBSum and store those in a matrix in memory
Once more, cut and export the monthly RGBSum for each ZOI and store them in a matrix where Matrix[[1]] is a list of monthly RGBSum for ZOI 1 and Matrix[[1]][[1]] is the the first month of the agricultural Year X for ZOI 1 (assuming entry data is 12 months and sorted by Agricultural Year) using as arguments: **RGBSum_YearX_ZOI** from step 28, **ZOI_Shapefile_List** from step 15, **ExportPath_RGBSum_Individual_YearX** from step 11, **ID_list** (global variable that should be created from running step 12), a string name for the files (XXX_, as the function will attach ZOI ID and month to the end), a string to print as each composite is exported, a string to print as each ZOI is finished and a string to print in the end;
```{r, eval=FALSE}
RGBSum_YearX_IndividualZOI_ListOfLists = IndividualZOIProduct(RGBSum_YearX_ZOI, ZOI_Shapefile_List, ExportPath_RGBSum_Individual_YearX, ID_list, "RGBSum_YearX_ZOI_", "Singular RGBSum Exported ", "Completed RGBSum ", "Completed All RGBSum")
```
### **32** - Reclassify our RGBSums
Although we have the normal sum of the RGB bands, we will need to reclassify it using our **VectorRGBSum** from step 13. For this matter, you can use **ReclassRasterList()**:
```{r, eval=FALSE}
RGBSum_YearX_ZOI_Binary = ReclassRasterList(RGBSum_YearX_ZOI, VectorRGBSum, "Reclassfied RGBSum ", "Reclassified all RGBSum")
```
### **33** - Export our recently reclassified RGBSums
Naturally, we can use **ExportRasterList()**:
```{r, eval=FALSE}
ExportRasterList(RGBSum_YearX_ZOI_Binary, ExportPath_RGBSum_Binary_All_YearX, "RGBSum_Binary_All_YearX_", "Finished Exporting Binary RGBSum ", "Finished Exporting all Binary RGBSum")
```
### **34** - Create Individual Folders for Binary RGBSum
We will of course also need folders for each binary RGBSum (for each ZOI). Since we are exporting reclassified rasters, let's use **CreateIndividualFolders2()**:
```{r, eval=FALSE}
CreateIndividualFolders2(ExportPath_RGBSum_Individual_YearX, ID_list, RGBSum_Threshold)
```
### **35** - Cut and Export Individual ZOI Binary RGBSum and store those in a matrix in memory
Once more, cut and export the monthly Binary RGBSum for each ZOI and store them in a matrix where Matrix[[1]] is a list of monthly Binary RGBSum for ZOI 1 and Matrix[[1]][[1]] is the the first month of the agricultural Year X for ZOI 1 (assuming entry data is 12 months and sorted by Agricultural Year) using as arguments: **RGBSum_YearX_ZOI_Binary** from step 28, **ZOI_Shapefile_List** from step 15, **ExportPath_RGBSum_Binary_Individual_** from step 11, **ID_list** (global variable that should be created from running step 12), a string name for the files (XXX_, as the function will attach ZOI ID and month to the end), a string to print as each composite is exported, a string to print as each ZOI is finished and a string to print in the end;
```{r, eval=FALSE}
Binary_RGBSum_YearX_IndividualZOI_ListOfLists = IndividualZOIProduct(RGBSum_YearX_ZOI_Binary, ZOI_Shapefile_List, ExportPath_RGBSum_Binary_Individual_YearX, ID_list, "RGBSum_Binary_YearX_ZOI_", "Singular Binary RGBSum Exported ", "Completed Binary RGBSum ", "Completed All Binary RGBSum")
```
## Processing Algorithm with Binary NDVI and Binary RGBSum
### **36** - Calculate Algorithm 
Finally, calculate the algorithm with **NDVI_YearX_ZOI_Binary** and **RGBSum_YearX_ZOI_Binary** from steps 19 and 32:
```{r, eval=FALSE}
Algorithm_YearX_ZOI = CalculateAlgorithm(NDVI_YearX_ZOI_Binary,RGBSum_YearX_ZOI_Binary)
```
### **37** - Reclassify Algorithm
Our algorithm is now a raster with 0, 1 and 2. 0 means no pixel met the threholds of both NDVI or RGBSum. 1 means one of them met the threshold, and 2 means both NDVI and RGBSum met the threshold. We'll now reclassify the raster to [0 and 1] = 0 and [2] = 1. This way we know which pixels are more likely to be water.
```{r, eval=FALSE}
Algorithm_YearX_ZOI_Binary = ReclassRasterList(Algorithm_YearX_ZOI, VectorAlgorithm, "Reclassfied Algorithm ", "Reclassified all Algorithm")
```
### **38** - Export reclassified Algorithm
Using **ExportRasterList()**, we can export our result (for all ZOI together):
```{r, eval=FALSE}
ExportRasterList(Algorithm_YearX_ZOI_Binary, ExportPath_Algorithm_All_YearX, "Algorithm_Binary_All_YearX_", "Finished Exporting Binary Algorithm ", "Finished Exporting all Binary Algorithm")
```
### **39** - Create Individual Folders for Algorithm
Use **CreateIndividualFolders3()**, as we can name the folders according to the thresholds used.
```{r, eval=FALSE}
CreateIndividualFolders3(ExportPath_RGBSum_Individual_YearX, ID_list, NDVI_Threshold, RGBSum_Threshold)
```
### **40** - Cut and Export Individual ZOI Binary Algorithm and store those in a matrix in memory
One final time, cut and export the monthly Binary Algorithm for each ZOI and store them in a matrix where Matrix[[1]] is a list of monthly Binary Algorithm for ZOI 1 and Matrix[[1]][[1]] is the the first month of the agricultural Year X for ZOI 1 (assuming entry data is 12 months and sorted by Agricultural Year) using as arguments: **Algorithm_YearX_ZOI_Binary** from step 28, **ZOI_Shapefile_List** from step 15, **ExportPath_Algorithm_Individual_** from step 11, **ID_list** (global variable that should be created from running step 12), a string name for the files (XXX_, as the function will attach ZOI ID and month to the end), a string to print as each composite is exported, a string to print as each ZOI is finished and a string to print in the end;
```{r, eval=FALSE}
Binary_Algorithm_YearX_IndividualZOI_ListOfLists = IndividualZOIProduct(Algorithm_YearX_ZOI_Binary, ZOI_Shapefile_List, ExportPath_Algorithm_Individual_YearX, ID_list, "Algorithm_Binary_YearX_ZOI_", "Singular Binary Algorithm Exported ", "Completed Binary Algorithm ", "Completed All Binary Algorithm")
```
## Building Dataframes for Time Series Graphs (and exporting .csv)
### **41** - Organize your Data into lists;
This list organization provides us with some more structuring. This is more complicated than it needs to be but it was a foundation for multiple thresholds at once. That being said, we will need:
<ol>
1. A list for all algorithms with threshold Y (Threshold Y = TY) for the year "YearX";
2. A list for all algorithms with all thresholds for the year "YearX";
</ol>
Naturally, list n1 goes inside list n2. What we want to store is obviously the Algorith, so use **Algorithm_YearX_ZOI_Binary**
```{r, eval=FALSE}
Binary_Algorithm_YearX_List = list()
Binary_Algorithm_YearX_TY_List = list()
Binary_Algorithm_YearX_TY_List[[1]] = Algorithm_YearX_ZOI_Binary
Binary_Algorithm_YearX_List[[1]] = Binary_Algorithm_YearX_TY_List
```
### **42** - Generate and export some .csv for further use. This will also return a Dataframe
**GenerateCSV2** will generate and export .csv for each ZOI with useful information like number of Water Pixels, Total Pixels etc.
```{r, eval=FALSE}
Algorithm_ZOI_Dataframe_List_YearX = GenerateCSV2(Binary_Algorithm_YearX_IndividualZOI_ListOfLists, ExportPath_CSV_YearX, "Agricultural Year X", ID_list, "Algorithm", NDVI_Threshold[[1]], RGBSum_Threshold[[1]])
```
### **43** - [OPTIONAL] Generate and export some .csv for further use. This will also return a Dataframe
It's optional, but if you'd like to compare the performance of our Algorithm vs the use of just the NDVI with your chosen threshold, you should use **GenerateCSV** to generate and export .csv for each ZOI with useful information so that it can always be red later.
```{r, eval=FALSE}
BinaryNDVI_ZOI_Dataframe_List_YearX =  GenerateCSV(NDVI_Binary_YearX_IndividualZOI_ListofLists, ExportPath_CSV_YearX, "Agricultural Year X", ID_list, "BinaryNDVI", NDVI_Threshold[[1]])
```
### You should do Steps 14 (included) to 43 (included) for each year of data you have. Then, proceed to the next step.
### **44** - Organize your Dataframes into lists
Let's assume you have 6 years of data, 2017-2022. As such, you should have 6 different **Algorithm_ZOI_Dataframe_List_YearX**, one for each Year. Now let's organize those into list.
```{r, eval=FALSE}
Algorithm_ZOI_Dataframe_List_All = list()
Algorithm_ZOI_Dataframe_List_All[[1]] = Algorithm_ZOI_Dataframe_List_2017
Algorithm_ZOI_Dataframe_List_All[[2]] = Algorithm_ZOI_Dataframe_List_2018
Algorithm_ZOI_Dataframe_List_All[[3]] = Algorithm_ZOI_Dataframe_List_2019
Algorithm_ZOI_Dataframe_List_All[[4]] = Algorithm_ZOI_Dataframe_List_2020
Algorithm_ZOI_Dataframe_List_All[[5]] = Algorithm_ZOI_Dataframe_List_2021
Algorithm_ZOI_Dataframe_List_All[[6]] = Algorithm_ZOI_Dataframe_List_2022
```
### **45** - [OPTIONAL] If you've done step 43, do the same for each **BinaryNDVI_ZOI_Dataframe_List_YearX** you have.
```{r, eval=FALSE}
BinaryNDVI_ZOI_Dataframe_List_All = list()
BinaryNDVI_ZOI_Dataframe_List_All[[1]] = BinaryNDVI_ZOI_Dataframe_List_2017
BinaryNDVI_ZOI_Dataframe_List_All[[2]] = BinaryNDVI_ZOI_Dataframe_List_2018
BinaryNDVI_ZOI_Dataframe_List_All[[3]] = BinaryNDVI_ZOI_Dataframe_List_2019
BinaryNDVI_ZOI_Dataframe_List_All[[4]] = BinaryNDVI_ZOI_Dataframe_List_2020
BinaryNDVI_ZOI_Dataframe_List_All[[5]] = BinaryNDVI_ZOI_Dataframe_List_2021
BinaryNDVI_ZOI_Dataframe_List_All[[6]] = BinaryNDVI_ZOI_Dataframe_List_2022
```
### **46** - Re-organize our structure from Year-Based to ZOI-Based
We need our Data to be structured on a ZOI to ZOI basis. Right now we have lists of all ZOI for a single year. What we need is a list of all years for a single ZOI. We can use **Create_ZOI_AllYears_List()** for that. Simply:
```{r, eval=FALSE}
Algorithm_All_ZOI_ByYear = Create_ZOI_AllYears_List(Algorithm_ZOI_Dataframe_List_All)
```
### **47** - [OPTIONAL] Re-organize our structure from Year-Based to ZOI-Based
Of course, if you've done steps 43 and 45, do the same for your **BinaryNDVI_ZOI_Dataframe_List_All**
```{r, eval=FALSE}
BinaryNDVI_All_ZOI_ByYear = Create_ZOI_AllYears_List(BinaryNDVI_ZOI_Dataframe_List_All)
```
## Plot our Time-Series
### **48** - Finally, it's time to plot. 
Simply use **PlotAllYears_All_ZOIs()** with your **Algorithm_All_ZOI_ByYear** and your **ExportPath_Graphs_Individual**
```{r, eval=FALSE}
PlotAllYears_All_ZOIs(Algorithm_All_ZOI_ByYear,ExportPath_Graphs_Individual)
```
### **49** - [OPTIONAL] If you've done steps 43, 45 and 47, you can now produce overlapping graphs to compare the performance of our Algorithm vs the performance of just the Binary NDVI. 
Use **PlotAllYears_All_ZOIs2()** with both your **BinaryNDVI_All_ZOI_ByYear**, your **Algorithm_All_ZOI_ByYear** and your **ExportPath_Graphs_Overlap_Individual**;
```{r, eval=FALSE}
PlotAllYears_All_ZOIs2(BinaryNDVI_All_ZOI_ByYear, Algorithm_All_ZOI_ByYear, ExportPath_Graphs_Overlap_Individual)
```
